---
title: "Creating and Applying Functions"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This document will help you to;

1. Create functions
2. Run your functions over, rows and/or columns, dataframes, lists etc. 

### 1. Defining a function

Give a working example of a function 

Let's say that we're working with a a dataframe, that consists of survey data
for a given population. 

Let's use the sampleSurvey data within the biomarkerstats package

```{r installation}
# install.packages("devtools")
# devtools::install_github("micronutrientsupport/biomarkerstats")
```

```{r sampleSurvey}
data <- biomarkerstats::sampleSurvey
head(data)
```

#### Create a function that applies to one row of the dataframe 

Let's say we'd like to create a function that does .... deficiency, simple? ...
for each individual.

Since we're going to be using the apply functions (base r, give link), we are 
going to right a function that can be 'applied' to each row (individual).

```{r function definition}
# insert function here...
```

Let's test it on one row of data (one individual in our survey dataset)

```{r row-test}

# Select one row of data
oneRow <- biomarkerstats::sampleSurvey[1, ]


# func <- function (x){
# perform operation
# perform operation
# perform operation
#return (x)
#}

# Invoke the function 
# func(oneRow)
```

## Apply the function over our dataset

Now let's 'apply' our function over our entire survey data 

Here's another example of a commonly used function in base r; mean ()

apply (x, 1, function) = applies function to rows  
apply (x, 2, function) = apply function to columns

Simply put: 

1 = apply to rows, 
2 = apply to columns 


apply(data["zinc"], 1, mean) = apply mean function to rows,  
!(edited for zinc, column only)

and collapse the other dimension (will get x numbers, as there's x rows) 

apply (data, 2, mean) = apply mean function to columns, and collapse the other
dimension (will get x numbers, as there's x columns) 


```{r apply-function}
# apply (data, 1, mean) 

```

### 3. What if my dataset is not a list 
Give examples of mapply, apply, also how to use split! (aggregate by different regions
and then )

```{r other apply functions}
# We are going to aggregate by region name
# Creates a list of data seperated by region name
region <- split(data, data$regionName)

# lapply output as a list (remove NAs)
lapply(region, function(x) colMeans(x[,c("ageInMonths","zinc")], na.rm=T))

# sapply output as a dataframe (remove NAs)
sapply(region, function(x) colMeans(x[,c("ageInMonths","zinc")], na.rm=T))

```

### 4. Apply the function over multiple datasets 

Embed 'v2_test_incl_urbanity' source code here. 

